---
title: "GHI Estimation using DLM"
output: html_document
---

# Data Pre-Processing

Let's clean the given data first, get rid of unnecessary hour values, we're only interested in GHI values for hour 8-16.

```{r}
df <- read.csv(file = './Bhadla new.csv')
temp = subset(df, Hour>7 & Hour<17)
ghi = temp[,c("Year","Month","Day","Hour","GHI","Clearsky.GHI")] # "Minute" not needed
head(ghi) # Check it's all correct
```

FIXME: Split the data 80:20 into training and testing data. (2001-2021)

```{r}
ghi$Date = as.Date(paste(ghi$Year, ghi$Month, ghi$Day, sep = "-"))
ghi$Date = as.POSIXct(paste(ghi$Date, ghi$Hour), format = "%Y-%m-%d %H")
ghi = ghi[,c("Date", "GHI", "Clearsky.GHI")]
require('dplyr')
train <- sample_frac(ghi, 0.8) 
sid  <- as.numeric(rownames(train))
test  <- ghi[-sid,]
head(train)
head(test)
```

Looks good, let's convert this into a time series object now.
Conversion of this dataframe slice into an xts object will allow us to use the relevant functionalities.

```{r}
require('xts')
traints <- xts(train$GHI, train$Date)
T = length(traints)
tail(traints) # Why does it still go till 2021 end?
plot(traints[1:T], main = "GHI", xlab = "Date", ylab = "GHI", major.format = "%d-%m-%Y %H")
# plot(ghits$Clearsky.GHI, main = "Clearsky GHI", xlab = "Date", ylab = "Clearsky GHI")
```

# First Attempt

Let's try to create a simple first order polynomial model.

```{r}
library(dlm)
index  <- 1:T # FIXME:  Correct x-label

model = dlmModPoly(order=1, dV=1, dW=1, m0=600, C0=500) # + dlmModSeas(12)
results_filtered_dlm = dlmFilter(traints[1:T], model)
results_smoothed_dlm = dlmSmooth(results_filtered_dlm)

plot(index, traints[1:T], ylab = "GHI", 
     main = "GHI",
     type='l', xlab="Datapoint",lty=3)
# points(index,traints[1:T],pch=1)
lines(index,results_filtered_dlm$m[-1],col='red',lwd=2)
lines(index,results_smoothed_dlm$s[-1],col='blue',lwd=2)
legend('topright', legend=c("Filtered", "Smoothened"),
       col = c("red","blue"), lty=c(1,1))
```

FIXME: Now, to use this trained model to check for test data. (What metrics?)

```{r}
testts <- xts(test$GHI, test$Date)
T2 = length(testts)
results_forecasted_dlm = dlmForecast(results_filtered_dlm, nAhead = T2)

index2 <-  1:T2
plot(index2, testts[1:T2], ylab = "GHI", 
     main = "GHI FC",
     type='l', xlab="Datapoint",lty=3,ylim=c(100,1000))
points(index2,testts[1:T2],pch=1)
lines(index2,results_forecasted_dlm$a[,1],col='red',lwd=2)
# lines(index2,results_smoothed_dlm$s[-1],col='blue',lwd=2)
# legend('topright', legend=c("Filtered", "Smoothened"),
#       col = c("red","blue"), lty=c(1,1))
length(traints)
length(testts)
```
